---
title: "Fishers Example"
author: "Simon Halliday"
date: "2015-July -19"
output: html_document
runtime: shiny
---

```{r, echo = FALSE}
require(shiny)
```


###This is stolen from the LaTex file and edited a bit
We assume the interaction between the two fishermen works as follows.
If Alfredo were fishing the lake by himself, and Bob were not there,
the amount of fish he would catch, $y^{A}$, when he fishes $e^{A}$
hours would be $12$ fish per hour::

$$y^{A}=12e^{A}$$


Alfredo, however, dislikes fishing, and deducts $\frac{1}{2}(e^{A})^{2}$
when he fishes $e^{A}$ hours, so his utility, or payoff, when he
fishes $e^{A}$ hours is

$$u^{A}=12e^{A}-\frac{1}{2}(e^{A})^{2}$$

Alfredo, fishing alone, would maximize utility by fishing $12$ hours
a day; his utility would be $72$. This would be Alfredo's \emph{decision
problem} if he had no interaction with Bob.

But Bob's fishing reduces the amount of fish Alfredo can catch in
an hour. For each hour Bob fishes, Alfredo catches $\frac{1}{2}$
fish per hour less. If Bob fishes $e^{B}$ hours, and Alfredo fishes
$e^{A}$ hours, Alfredo's catch is:

$$y^{A}=(12-\frac{1}{2}e^{B})e^{A}$$


Thus the social interaction in the situation is that Alfredo's utility
depends on both his own hours of fishing, and on Bob's hours of fishing:

$$u^{A}=(12-\frac{1}{2}e^{B})e^{A}-\frac{1}{2}(e^{A})^{2}$$


We assume that Bob has the same utility dependent on Alfredo's fishing.

To keep things simple, suppose for the moment that the only actions
Alfredo and Bob consider are to fish either $6$ or $8$ hours.

The two fishermen's actions lead to the following four possible outcomes: 

- **Alfredo fishes 8 hours, Bob fishes 8 hours**: When both fishermen
fish 8 hours, we can see that each one catches $64$ fish and has
a utility of $32$. Each fisherman ends up with the same payoff as
his fallback position, $32$. 
-  **Alfredo fishes 6 hours, Bob fishes 6 hours**: When both fishermen
spend less time fishing they each catch more fish per hour. They also
benefit because they are spending less time fishing. Each fisherman
catches $54$ fish and has a utility of $36$, which is larger than
their fallback position, $32$. 
- **Alfredo fishes 6 hours, Bob fishes 8 hours**: Because Bob fishes
8 hours, In this case Bob benefits because with Alfredo fishing less,
he catches more fish per hour. Bob catches $72$ fish and has a utility
of $40$. Alfredo, however, is catching less fish per hour and fishing
less; he catchs $48$ fish and has a utility of $30$, which is less
than the fallback position $32$.
- **Alfredo fishes 8 hours, Bob fishes 6 hours**: This is symmetrical
to the previous description, so now Alfredo's utility is $40$, and
Bob's is $30$.

We can take the description of the players, actions and payoffs we
provided above and convert it into a *normal form game* as represented
by th *payoff matrix* in the Table. 
The convention we will use throughout this book it to list the row
player's payoffs first and the column player's payoffs second. So,
in the Fishermen's Dilemma game, Alfredo's payoffs are first and Bob's
payoffs are listed second. 

|   |   | Bob | |
|---|---|---|---|
|   |   | Fish 6 Hours | Fish 8 Hours |
| Alfredo | Fish 6 Hours | 36, 36 | 30, 40 |
|  | Fish 8 Hours | 40, 30 | 32, 32 |


Here we describe the problem of the fishers dilemma in Ch. 2 or Ch. 8. 
We can see below that we have parameterized the model in the same way that we did 
in chapter 2 of the textbook in the one Box. 

perhaps we want some math, as we have in the textbook: 

##The fishers' best response functions
We can also find the best repsonses by differentiating $u^A$ with respect to $e^A$ and setting the result equal to zero to find the utility-maximizing level of effort for A, which gives the first-order condition: 
$$\frac{\partial u^A}{\partial e^A} = \alpha(e - \beta e^B) - e^A = 0$$
The FOC implies that: 
$$e^A[e^B] = \alpha(1 - \beta e^B)$$
$$e^B[e^A] = \alpha(1 - \beta e^A)$$

Let us assume that $\alpha$ and $\beta$ start at 12 and 0.5 respectively, but that they might actually vary for each player. These ideas are captured in the simulation below.

##Graphs of the Fishers' Best Responses
```{r, echo = FALSE}
shinyApp(
  ui = fluidPage(
   sidebarLayout(
    sidebarPanel(
                sliderInput("T",
                                HTML("Fisher A: amount of fish per hour"),
                                min = 1,
                                max = 20,
                                value = 12,
                                step = 1), 
                    
                sliderInput("m",
                                HTML("Fisher A: amount of externality"),
                                min = 0,
                                max = 1,
                                value = 0.5,
                                step=0.1,
                                round = FALSE),
                sliderInput("Q",
                                HTML("Fisher B: amount of fish per hour"),
                                min = 1,
                                max = 20,
                                value = 12,
                                step = 1), 
                sliderInput("s",
                                HTML("Fisher B: amount of externality"),
                                min = 0,
                                max = 1,
                                value = 0.5,
                                step=0.1,
                                round = FALSE)
   ),
 mainPanel(plotOutput("BestResponse")
             )
     )
  ),
server = function(input, output) {
  output$BestResponse <- renderPlot(width = 400, {
    BestResponse1 <- function(a, T=12, m=0.5) {
      ((T/m)-(1/m)*a)
    }
    BestResponse2 <- function(a, Q=12, s=0.5){
      (Q-s*a)
    }
    COL <- c("#1B9E77", "#F0F0D8", "#D8D8C0", "#E7298A", "#66A61E", "#E6AB02", "#A6761D", "#666666")
    par(mar =  c(2, 2, 5, 5))
    xlims <- c(0, 30)
    ylims <- c(0, 30)
    
    plot(0, 0, xlim = xlims, ylim = ylims, type = "n",
         xlab = expression(paste("amount of A, ", a)),
         ylab = expression(paste("amount of B, ", b)),
         xaxt = "n", 
         yaxt = "n", 
         cex.lab = 1.5, 
         bty = "n")
    
    npts <- 500
    xx1 <- seq(xlims[1], input$T, length.out = npts)
    yy1 <- BestResponse1(xx1, T=input$T, m=input$m)
    lines(xx1, yy1, col = COL[1], lwd = 4)
    xx2 <- seq(xlims[1], input$Q / input$s, length.out = npts)
    yy2 <- BestResponse2(xx2, Q=input$Q, s=input$s)
    lines(xx2, yy2, col = COL[3], lwd = 4)
    
    
    axis(1, at = xlims, pos = 0)
    axis(2, at = ylims, pos = 0)
    add_legend <- function(...) {
      opar <- par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), 
                  mar=c(0, 0, 0, 0), new=TRUE)
      on.exit(par(opar))
      plot(0, 0, type='n', bty='n', xaxt='n', yaxt='n')
      legend(...)
    }
    
    add_legend("topright", legend=c(
      expression(paste("A best response function")), 
      expression(paste("B best response function"))
    ),
    col = COL, lty = 1, lwd = 3, bty = "n", cex=1.2, pt.cex = 1)
  })
}, 
  options = list(height = 500)
)
```




