---
title: "CVC Workshop Notes"
author: "Simon Halliday"
date: "2015-July -13"
output: 
  html_document:
    css: ../stylesheets/styles.css

---


Importing the medicare data... 
```{r}
#medicare <- read.csv("/Users/shalliday/Google Drive/simondhalliday.github.io/cvc_workshop/medicare_fy2013.csv")
medicare <- read.csv("medicare_fy2013.csv")
#head(medicare)
#str(medicare)
```

Trying to embed a badly thought-out plot using `ggvis`:

```{r}
library(dplyr)
library(ggvis)
medicare %>% 
  ggvis(~factor(Provider.State), ~Average.Covered.Charges) %>%
  layer_bars()

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



##Teaching Tips
These are tips I picked out (thus peculiar to my tastes)

* __For teaching the idea of the command line and the script:__ How many of you had a [TI calculator](https://en.wikipedia.org/wiki/Comparison_of_Texas_Instruments_graphing_calculators)? (TI = Texas Instruments) Did you know how to use it on the very first day? What about on your final day of high school? 

* When R outputs the error `object '[object name]' not found` that normally means you (or your student) have mis-spelt the variable name or dataframe name. 


