---
title: "CVC Workshop Notes"
author: "Simon Halliday"
date: "2015-July -13"
output: 
  html_document:
    css: ../stylesheets/styles.css

---


Importing the medicare data... 
```{r}
#medicare <- read.csv("/Users/shalliday/Google Drive/simondhalliday.github.io/cvc_workshop/medicare_fy2013.csv")
medicare <- read.csv("medicare_fy2013.csv")
#head(medicare)
#str(medicare)
```

Trying to embed a badly thought-out plot using the `ggvis` package:

```{r}
library(dplyr)
library(ggvis)
medicare %>% 
  ggvis(~factor(Provider.State), ~Average.Covered.Charges) %>%
  layer_bars()

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



##Teaching Tips
These are tips I picked out (thus peculiar to my tastes)

* __For teaching the idea of the command line and the script:__ How many of you had a [TI calculator](https://en.wikipedia.org/wiki/Comparison_of_Texas_Instruments_graphing_calculators)? (TI = Texas Instruments) Did you know how to use it on the very first day? What about on your final day of high school? (RP)

* When R outputs the error `object '[object name]' not found` that normally means you (or your student) have mis-spelt the variable name or dataframe name. (RP)

* Pick your favorite way (command) for doing a job, don't show students too many methods. You'll overwhelm them if you give them too much. __Less volument, more creativity__. (RP)

* Call tildes `~` 'wiggles'! Sounds funny and is less scary then tilde. (RP) 

* Two most important questions: What do you want R to do? (goal) What must R know to do that? (RP) [me - why don't I call R a specific name? e.g. R = Rachel - What do I want Rachel to do? What must Rachel know to do that?]

## Commonplace Book

* RP: "If your students find R to be the hardest part of your course, your R is too hard and your questions are too easy."


## Commands to run
Scatterplot with birth data:
```{r}
library(mosaic)
library(mosaicData)
xyplot(births ~ dayofyear, data = Births78)

```


```{r}

```



