---
title: "SAIRR Check"
author: "Simon Halliday"
date: "2017-June -25"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warnings = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
library(stargazer)
```

Import the data from the GHS 2015:
```{r data_import, warnings = FALSE, message = FALSE}
ghs15 <- read_csv("uct/csv/ghs_2015_person_v1.1_20160608.csv")
```

##Wrangling 
Wrangle the data and select the sample of 20 and older. 
Generate a variable that checks the proportion of people of each race that has Matric or higher. 
Also try to find out their error with another variable. 
Couldn't quite get the same proportions they did, but I wasn't trying too hard. 
```{r}
educ20 <- 
  ghs15 %>%
  filter(Age > 19) %>%
  mutate(matricmore = ifelse(Q15HIEDU %in% c(12, 13, 16,
                                             17, 18, 19,
                                             21, 22, 23,
                                             24, 25, 26,
                                             27, 28, 29),
                             1, 0), 
         matriconly = ifelse(Q15HIEDU %in% c(12, 13, 23)
                             , 1, 0), 
         race = factor(race, labels = c("Black/African",
                                        "Coloured",
                                        "Indian/Asian",
                                        "White")))
```

Generate summary data frame. 
```{r}
raceSum <- 
  educ20 %>%
  group_by(race) %>%
  summarise(matricprop = round(mean(matricmore), 3))
stargazer(raceSum, type = "text", summary = FALSE)
```

Screwing around with some other tallies to check my work above. 
```{r}
mosaic::tally(~Q15HIEDU, data = educ20)
mosaic::tally(~race, data = educ20)
mosaic::tally(~matricmore, data = educ20)
mosaic::tally(~matricmore | race, data = educ20, format = "percent")
```


