---
title: "Quiz 1"
author: "Simon Halliday"
date: "2015-September -24"
output: html_document
---

Load the packages you require: `mosaicData` and `DataComputing`: 
```{r requirements, message = FALSE}
library(mosaicData)
library(DataComputing)
```


I picked a name randomly from the class and got "Riley". We shall investigate a little bit about that name using the `BabyNames` datatable provided in `mosaicData`. 

I ran the following command to create a data table called `Riley` to
provide a table with the total of all people called Riley in the `BabyNames` data table.

Copy and paste this command into your console or R markdown in a data
chunk and run it. 
```{r Riley}
Riley <- 
  BabyNames %>%
  filter(name == "Riley") %>%
  group_by(year, sex) %>%
  summarise(yearlytotal = sum(count))
```

You should now have an object in your "Environment" tab called "Riley". 

For the question in the quiz, you need to use the function `head()` to find out how many people are called Riley at the top of the data table Riley.
```{r}
head(Riley)
```

Now, we need to `filter()` this table to create a new data table (I called mine `FemRiley`) to get a data table that includes only the female cases called Riley. 
```{r FemRiley, include = FALSE}
FemRiley <- 
  Riley %>% 
  filter(sex == "F")
FemRiley
```

Now, for the question in the quiz, you need to find out the dimensions of your DataTable. To find out the dimensions of a data table we use the function `dim()`:
```{r dimFemRiley, results = "hide"}
dim(FemRiley)
```

For the next question in the quiz, you need to calculate the number of
females called Riley in the 1995 (I don't know what year Riley at Smith
was born, but I took an educated guess. Riley, apologies if I got it
incorrect.) Once more, use filter to create an object called FemRiley1995
with only those people called Riley and born in 1995 in the object
FemRiley1995. You will need to use your insight from the class to do this. 

```{r, include = FALSE, results = "hide"}
FemRiley1995 <- 
  FemRiley %>% 
  filter(year == 1995)
```

```{r, results = "hide"}
FemRiley1995
```



